\chapter{Introduction}
\label{sec:introduction}
%\thispagestyle{empty}
\section{Learning dictionaries}

%Learning the most significant basis elements, that make up an
%signal, is a popular task in signal processing.
\Todo{write at the end, merge with overview?}
%Signal <> is a topic <> since the 1960s. The basic theory is <>. Over the last
%few decades it has been
%a major <> to develop and find new bases for the signals. In the last 15 years 
%\Todo{get the reader}
Compressed sensing .. Data networks, data transfer, error correction. 

Experiments \cite{Chen1998,Elad2006} have shown that just the right basis
directly learned from images can yield much better results in reconstruction
than those based on mathematical functions such as cosine transforms or
wavelet transforms. These learned basis elements can lead to better
reconstruction quality of images in large image databases and assist tasks such
as classification, noise removal, image enhancement, organization and
compression.

Learning such basis is often limited to small training sets of 100.000 to
1.000.000 samples. But large databases have training sets of millions and more.
In 2009 Mairal et al.\cite{Mairal2009} presented an new online learning
algorithm that can constantly consume new samples rather than just a fixed
batch.  We evaluate this algorithm to 
Can large databases benefit from that? Can we effectivly learn dictionaries for
large databases? What else can we learn from these dictionaries?

As an example application to use these dictionaries we present a simple 
compression algorithm. We evaluation sparse coding and basis learning algorithms
and strategies for large image databases and transport recent discoveries in
sparse coding and basis learning into a real world example of image compression.

\section{Structure of thesis}
The next chapter gives an overview on the idea of sparse signal
representation with a short excurs into the history of signal analysis.
Chapter three describes the fundamental theory of sparse coding, introduces the
most common algorithms and presents usual applications of sparse coding. The
fourth chapter introduces algorithms for the design of over-complete
dictionaries of basis elements and presents current related work in this field.
The fifth chapter describes the applied methods for sparse
coding, the generation of dictionaries via clustered training with a large image
database of 100,000 to 1,000,000 images collected from \url{flickr.com}.
Presents experiments regarding the structure of dictionary elements, convergence
of the training process, comparison of signal representation quality and
adaptability of sparse coding for image compression. The two final chapters
present the results of the experiments, discuss these results and new findings
and show potential improvements in future work. 

\section*{Notation}
\begin{tabular}{c l}
$x$ & scalar\\
$T$ & constant\\
$\vec{x}$ & vector\\
$\vec{1}$ & vector with all elements 1\\
$\vec{x}[i]$ & $i$-th element of vector $\vec{x}$\\
$\mat{M}$ & matrix\\
$A$ & set\\
$A^C$ & complment of set $A$ in relation to universal set $U$\\ 
$\vec{x}_A$ & vector $\vec{x}$ reduced to rows of set $A$ \\
$\mat{M}_A$ & matrix $\mat{M}$ reduced to rows and columns of set $A$ \\

\end{tabular}



